### 2015-05-12 whoop.tex
more fixes

$\mutexunlock{m};$ & mutex unlock, where $m \in \mutexes$\\
\bottomrule
\end{tabular}
\caption{Forms of statements in our simple programming model}\label{fig:statements}
\end{figure}

@@

$\mutexunlock{m};$ & $\currentlockset{i} = \currentlockset{i} \setminus \{ m \};$ \\
\bottomrule
\end{tabular}
\caption{Instrumenting statements for lockset analysis}\label{fig:instrumentation}
\end{figure}

@@

\end{tabular}
\caption{The sequential program to be analysed in order to prove race-freedom for a pair of thread templates}
\label{fig:sequentialization}
\end{figure}

\medskip\noindent\textbf{Soundness }
%
We sketch an argument that if the program of Figure~\ref{fig:sequentialization} is correct,
it is impossible for a thread executing template $T$ to race with a thread executing template $U$,
under the assumption that the threads are guaranteed to terminate.
Let us assume that the program of Figure~\ref{fig:sequentialization} is correct,
and suppose (by way of contradition) that a thread executing $T$ can in fact race with a thread
executing $U$, on some shared variable $s$.  By our hypothesis that the program
of Figure~\ref{fig:sequentialization} is correct, and that the threads terminate,
the assertion checked at the end of the program guarantees at least one mutex, $m$ say,
belongs to both $\lockset{s}{i}$ and $\lockset{s}{j}$.  By the definition of a lockset
(and according to the manner in which shared accesses are instrumented in Figure~\ref{ }),
this means that $m$ is held during every access to $s$ by both $\instrument{T}{i}$ and $\instrument{U}{j}$.
As a result, $m$ must be unlocked and locked between the two accesses, which
contradicts that the pair of accesses is racing.

We require the threads to terminate becauase in the presence of non-termination the assertion
at the end of Figure~\ref{fig:sequentialization} may not be reached.  The termination analysis
problem for device drivers has been widely studied, notably in the context of the Terminator tool~\cite{cook2006termination}.
In the remainder of the paper we do not consider termination issues, assuming that the drivers we analyse
in our experimental evaluation (Section~\ref{evaluation}) are terminating.

>>>>

$\mutexunlock{m};$ & mutex unlock, where $m \in \mutexes$\\
\bottomrule
\end{tabular}
\caption{Forms of statements in our simple programming model}
\label{fig:statements}
\end{figure}

@@

$\mutexunlock{m};$ & $\currentlockset{i} = \currentlockset{i} \setminus \{ m \};$ \\
\bottomrule
\end{tabular}
\caption{Instrumenting statements for lockset analysis}
\label{fig:instrumentation}
\end{figure}

@@

\end{tabular}
\caption{The sequential program to be analyzed in order to prove race-freedom for a pair of thread templates}
\label{fig:sequentialization}
\end{figure}

\medskip\noindent\textbf{Soundness }
%
We sketch an argument that if the program of Figure~\ref{fig:sequentialization} is correct, it is impossible for a thread executing template $T$ to race with a thread executing template $U$, under the assumption that the threads are guaranteed to terminate. Let us assume that the program of Figure~\ref{fig:sequentialization} is correct, and suppose (by way of contradiction) that a thread executing $T$ can in fact race with a thread executing $U$, on some shared variable $s$.  By our hypothesis that the program of Figure~\ref{fig:sequentialization} is correct, and that the threads terminate, the assertion checked at the end of the program guarantees at least one mutex, say $m$, belongs to both $\lockset{s}{i}$ and $\lockset{s}{j}$.  By the definition of a lockset (and according to the manner in which shared accesses are instrumented in Figure~\ref{fig:instrumentation}), this means that $m$ is held during every access to $s$ by both $\instrument{T}{i}$ and $\instrument{U}{j}$. As a result, $m$ must be unlocked and locked between the two accesses, which contradicts that the pair of accesses is racing.

We require the threads to terminate because in the presence of non-termination the assertion at the end of Figure~\ref{fig:sequentialization} may not be reached.  The termination analysis problem for device drivers has been widely studied, notably in the context of the Terminator tool~\cite{cook2006termination}. In the remainder of the paper we do not consider termination issues, assuming that the drivers we analyze in our experimental evaluation (see Section~\ref{evaluation}) are indeed terminating.
